### 2013-02-14 doc/xps2013/proposal.tex
edits to summary

Grappa applications are likely to use large amounts of data, 100s of terabytes to petabytes. Keeping all these data in DRAM only is unlikely to be cost effective. We will explore mechanisms that use solid-state storage as primary memory, greatly expanding the directly addressible storage at low cost.

@@

Irregular applications are increasing in importance to the computing industry.  For a long time they have underpinned computations of interest to the national security arm of the federal government.  But now, they underpin such questions as ad placement in social networks, and analysis of complex data-sets in medicine and science.  The defining characteristic of these applications is poor locality and massive available parallelism.  Our effort is focused on making these applications perform well and scale on commodity hardware.  The ``secret sauce'' that makes this work is to essentially build a software XMT~\cite{feo:xmt}.  Instead of relying on locality (which doesn't exist for these applications) we rely on threading to tolerate memory latency.  Unlike the XMT, however, we must contend with commodity networking hardware, and here we again rely on threading to provide a sufficient number of requests that we can buffer them prior to dispatching on the physical wire.

Our past effort has shown a workable proof of concept of all this runtime smoke and mirrors: a commodity cluster can out do the XMT at its own game, by emulating its key features in software.  What is needed now is additional effort to improve the underlying performance of this runtime and make it usable to a broad base of developers.  We will be focusing our efforts in the next three years on the key performance component, the networking layer, and the key programmer efficiency linchpin, high level language support.  In addition we need a useful mechanism for handling data-sets in the petabyte range, and here we imagine building a cluster of machines that relies primarily on SSD's and not DRAM for data-set storage.

>>>>

Grappa applications are likely to use large amounts of data, 100s of terabytes to petabytes. Keeping all these data in DRAM only is unlikely to be cost effective. We will explore mechanisms that use solid-state storage as primary memory, greatly expanding the directly addressable storage at low cost.

@@

Irregular applications are increasing in importance to the computing industry.
For a long time they have underpinned computations of interest to the national
security arm of the federal government. But now, they underpin such questions
as ad placement in social networks, and analysis of complex data-sets in
medicine and science. The defining characteristic of these applications is
poor locality and massive available parallelism. Our effort is focused on
making these applications perform well and scale on commodity hardware. The
key idea that makes this work is to rely on massive concurrency to tolerate
memory latency, instead of relying on locality (which doesn't exist for these
applications). This core idea underpinned the XMT~\cite{feo:xmt} system, a
fully-custom hardware system. We plan to explore that idea in software only,
running on commodity processors. And unlike the XMT, however, we must also
contend with commodity networking hardware, and here we again rely on
threading to provide a sufficient number of requests that we can buffer them
prior to dispatching on the network, to overcome bandwidth limitations when
operating with small messages.

Our preliminary effort has shown a workable proof of concept of all this
runtime smoke and mirrors: a commodity cluster can out do the XMT at its own
game, by emulating its key features in software. What is needed now is
additional effort to improve the underlying performance of this runtime and
make it usable to a broad base of users. We will be focusing our efforts
in the next three years on the key performance component, the networking
layer, and the key programmer efficiency linchpin, high level language
support. In addition we need a useful mechanism for handling data-sets in the
petabyte range, and here we imagine building a cluster of machines that relies
primarily on SSD's and not DRAM as main memory data-set storage.
